---
title: "Data Ingestion/EDA"
author: "Emilio Sanchez San Martin"
date: "2025-04-06"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

## Continuing Data Investigation

First, I have to get every library that I will use for this Data
Investigation/EDA.

```{r}
library(tidyverse)
library(ggplot2)
```

I have to put the three datasets now on my global envrionment to work
with them.

```{r}
# Loading the datasets
High_Volume_Weekly_DS <- read.csv("/Users/emilio/Downloads/Montgomery-College-Data-Set-1(High-Volume-Weekly)-csv.csv")
Medium_Volume_Weekly_DS <- read.csv("/Users/emilio/Downloads/Montgomery-College-Data-Set-1(Medium-Volume-Weekly)-csv.csv")
Low_Volume_Weekly_DS <- read.csv("/Users/emilio/Downloads/Montgomery-College-Data-Set-1(Low-Volume-Weekly)-csv.csv")

```

I noticed that the variables on the top of my data set have X1, X2, X3,
etc. I will remove the X and replace it with "Week".

```{r}
# Renaming the columns
colnames(High_Volume_Weekly_DS) <- gsub("X", "Week", colnames(High_Volume_Weekly_DS))
colnames(Medium_Volume_Weekly_DS) <- gsub("X", "Week", colnames(Medium_Volume_Weekly_DS))
colnames(Low_Volume_Weekly_DS) <- gsub("X", "Week", colnames(Low_Volume_Weekly_DS))
```

Finally, I will do my last step with this data set, which is to multiply
to "Bottles_Per_Case" varibale with "Cost_Amount_Per_Bottle" variable to
get the "Total_Cost" variable for all the bottles in a case. I will do
this for all three data sets.

```{r}
# Creating the Total_Cost variable
High_Volume_Weekly_DS <- High_Volume_Weekly_DS |>
  mutate(Total_Cost = Bottles_Per_Case * Cost_Amount_Per_Bottle)

Medium_Volume_Weekly_DS <- Medium_Volume_Weekly_DS |>
  mutate(Total_Cost = Bottles_Per_Case * Cost_Amount_Per_Bottle)

Low_Volume_Weekly_DS <- Low_Volume_Weekly_DS |>
  mutate(Total_Cost = Bottles_Per_Case * Cost_Amount_Per_Bottle)

```

I want to move the last column "Total_Cost" right after the
"Cost_Amount_Per_Bottle" column. I will do this for all three data sets.

```{r}
# Moving the Total_Cost column
High_Volume_Weekly_DS <- High_Volume_Weekly_DS |>
  select(ItemID, Description, Bottles_Per_Case, Cost_Amount_Per_Bottle, Total_Cost, everything())
Medium_Volume_Weekly_DS <- Medium_Volume_Weekly_DS |>
  select(ItemID, Description, Bottles_Per_Case, Cost_Amount_Per_Bottle, Total_Cost, everything())
Low_Volume_Weekly_DS <- Low_Volume_Weekly_DS |>
  select(ItemID, Description, Bottles_Per_Case, Cost_Amount_Per_Bottle, Total_Cost, everything())

```

Finished! I will do my last step, which is to check the structure of the
data sets to see if everything is in order. I will do this by using the
str() function.

```{r}
# Checking the structure of the datasets
str(High_Volume_Weekly_DS)
str(Medium_Volume_Weekly_DS)
str(Low_Volume_Weekly_DS)

```

Time to export the data sets to CSV files so I can use them in my
analysis.

```{r}
# Exporting the datasets to CSV files
write.csv(High_Volume_Weekly_DS, "/Users/emilio/Downloads/High_Volume_Weekly_DS.csv", row.names = FALSE)
write.csv(Medium_Volume_Weekly_DS, "/Users/emilio/Downloads/Medium_Volume_Weekly_DS.csv", row.names = FALSE)
write.csv(Low_Volume_Weekly_DS, "/Users/emilio/Downloads/Low_Volume_Weekly_DS.csv", row.names = FALSE)

```

## Now time to work on the EDA!

My next goal for this project is to discover patterns with the data, and
what I can find to help others know to understand. Particullarly, if I
were able to find Moving Average Sales trends from any product or
products, especially a top selling product, I can understand the basis
of how my algorithm could work. My first goal will try to uncover...

##### - Moving Average Salaes Trends for a Top-Selling Product

Working with Weekly data especially will help look at trends better.

I will choose the most sold amount of products across all stores, for
ALL of 2024.

Since we have the "Grand_Total" Variable, all I have to do is look at
the highest grand total for all the products and see which is the
highst.

```{r}
# Finding the top-selling product for 2024
top_selling_product_high_vol_store <- High_Volume_Weekly_DS |>
  arrange(desc(Grand_Total)) |>
  slice(1)

top_selling_product_medium_vol_store <- Medium_Volume_Weekly_DS |>
  arrange(desc(Grand_Total)) |>
  slice(1)

top_selling_product_low_vol_store <- Low_Volume_Weekly_DS |>
  arrange(desc(Grand_Total)) |>
  slice(1)

top_selling_product_high_vol_store
top_selling_product_medium_vol_store
top_selling_product_low_vol_store
```

As we see above, the top selling product for the.. High Volume store =
"SCOTTY'S VODKA 50ML" Medium Volume store = "FIREBALL CINN WHISKEY
50ML/10PK LOOSE" Low Volume store = "FIREBALL CINN WHISKEY 50ML/10PK
LOOSE".

Now for curiosity, I want to use the moving average across a set of
weeks to see if I can find trends. I will have to use the zoo package
for this.

```{r}
library(zoo) #Used Google to help understnad how to use this package

```

```{r}
# Sales for 53 weeks in a vector
weekly_sales_high <- as.numeric(High_Volume_Weekly_DS[1, paste0("Week", 1:53)])
weekly_sales_medium <- as.numeric(Medium_Volume_Weekly_DS[1, paste0("Week", 1:53)])
weekly_sales_low <- as.numeric(Low_Volume_Weekly_DS[1, paste0("Week", 1:53)])

# Will use a 4-week moving average
moving_avg_high <- rollmean(weekly_sales_high, k = 4, fill = NA, align = "right")
moving_avg_medium <- rollmean(weekly_sales_medium, k = 4, fill = NA, align = "right")
moving_avg_low <- rollmean(weekly_sales_low, k = 4, fill = NA, align = "right")

# Below I will put approximate starting week for each month to show when the months start
month_weeks <- c(1, 5, 9, 14, 18, 22, 27, 31, 36, 40, 45, 49)
month_labels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")


plot(moving_avg_high, type = "l", col = "darkblue", lwd = 2,
     xaxt = "n",  # Turning off x-axis to show months
     main = "4-Week Moving Average - High Volume Store",
     ylab = "Sales", xlab = "Week",
     cex.main = 1.2, cex.lab = 1.1, cex.axis = 0.9)
axis(1, at = month_weeks, labels = month_labels)
grid()

plot(moving_avg_medium, type = "l", col = "darkgreen", lwd = 2,
     xaxt = "n",  # Turning off x-axis to show months
     main = "4-Week Moving Average - Medium Volume Store",
     ylab = "Sales", xlab = "Week",
     cex.main = 1.2, cex.lab = 1.1, cex.axis = 0.9)
axis(1, at = month_weeks, labels = month_labels)
grid()

plot(moving_avg_low, type = "l", col = "darkred", lwd = 2,
     xaxt = "n",  # Turning off x-axis to show months
     main = "4-Week Moving Average - Low Volume Store",
     ylab = "Sales", xlab = "Week",
     cex.main = 1.2, cex.lab = 1.1, cex.axis = 0.9)
axis(1, at = month_weeks, labels = month_labels)
grid()
```

Oh wow.. as you can see from the above, the 4-week Moving Averages for
differen volume of stores for the HIGHEST selling product all year long
has been different for different weeks. Towards the end of the year, the
sales for the high volume store have been increasing, while the sales
for the medium and low volume stores have been decreasing in the middle
of the weeks (as you can see... around July and August). This is very
interesting to see. I will have to look at this more closely later on.
